---
title: "Intorduction to tabyls with NHIS data"
author: "Harrison Dekker"
date: '2022-04-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Processing NHIS data from IPUMS
IPUMS has a unique way of packaging data. This code snippet shows how to read a downloaded dataset.

```{r, eval = FALSE}
if (!require("ipumsr")) stop("Reading IPUMS data into R requires the ipumsr package. It can be installed using the following command: install.packages('ipumsr')")

ddi <- read_ipums_ddi("nhis_00007.xml")
data <- read_ipums_micro(ddi)
```

For our class, we will only select a few variables. I have prepared a dataset that we'll be using. The code below shows the steps involved.

```{r, eval=FALSE}
nhis <- df %>%
  mutate(REGION = as_factor(REGION),
         SEX = as_factor(SEX),
         EDUC = as_factor(EDUC),
         HEALTH = as_factor(HEALTH),
         HINOTCOVE = as_factor(HINOTCOVE),
         DEPFEELEVL = as_factor(DEPFEELEVL),
         CVDSYMP = as_factor(CVDSYMP)
  ) %>%
  select(YEAR, REGION, AGE, SEX, EDUC, HEALTH, HINOTCOVE, DEPFEELEVL, CVDSYMP, PERWEIGHT)

save(nhis, file="nhis-example.Rdata")
```

The data file `nhis-example.Rdata` is available on Brightspace. Download it now, then
create a new project in RStudio and load it as follows:

```{r}
load("original-data/nhis-example.Rdata")
View(nhis)
```

```
Codebook
YEAR                Year os survey

REGION              Region of residence
01                  Northeast
02                  North Central/Midwest
03                  South
04                  West
08                  NO DATA IN ROUND
09                  Unknown

SEX                 Sex
1                   Male
2                   Female
7                   Unknown-refused
8                   Unknown-not ascertained
9                   Unknown-don't know

EDUC                Educational attainment
000                 NIU
100                 Grade 12 or less, no high school diploma or equivalent
101                 Grade 8 or less (no further detail)
102                 Never attended/kindergarten only
103                 Grades 1-11 (no further detail)
104                 Grade 1
105                 Grade 2
106                 Grade 3
107                 Grade 4
108                 Grade 5
109                 Grade 6
110                 Grade 7
111                 Grade 8
112                 Grade 9-12, no diploma (no further detail)
113                 Grade 9
114                 Grade 10
115                 Grade 11
116                 12th grade, no diploma
200                 High school diploma or GED
201                 High school graduate
202                 GED or equivalent
300                 Some college, no 4yr degree
301                 Some college, no degree
302                 AA degree: technical/vocational/occupational
303                 AA degree: academic program
400                 Bachelor's degree (BA,AB,BS,BBA)
500                 Master's, Professional, or Doctoral Degree
501                 Master's degree (MA,MS,Med,MBA)
502                 Professional (MD,DDS,DVM,JD)
503                 Doctoral degree (PhD, EdD)
504                 Other degree
996                 No degree, years of education unknown
997                 Unknown--refused
998                 Unknown--not ascertained
999                 Unknown--don't know

HEALTH              Health status
1                   Excellent
2                   Very Good
3                   Good
4                   Fair
5                   Poor
7                   Unknown-refused
8                   Unknown-not ascertained
9                   Unknown-don't know

HINOTCOVE           Health Insurance coverage status
0                   NIU
1                   No, has coverage
2                   Yes, has no coverage
7                   Unknown-refused
8                   Unknown-not ascertained
9                   Unknown-don't know

DEPFEELEVL          Level of depression, last time depressed
0                   NIU
1                   A little
2                   A lot
3                   Somewhere between little and lot
7                   Unknown-refused
8                   Unknown-not ascertained
9                   Unknown-don't know

CVDSYMP             Severity of coronavirus symptoms at their worst
0                   NIU
1                   No symptoms
2                   Mild symptoms
3                   Moderate symptoms
4                   Severe symptoms
7                   Unknown-refused
8                   Unknown-not ascertained
9                   Unknown-don't know

```
